Against Spurious Understanding: Dovelating Inflated AI Claims \raisebox{-6pt}{\includegraphics[scale=0.09]{dove_1f54a-fe0f.png}}

### 2024-01-31 main.tex
Updates from Overleaf

\begin{icmlauthorlist}
\icmlauthor{Patrick Altmeyer}{yyy}
\icmlauthor{Antony Bartlett}{equal, yyy}
\icmlauthor{Andrew M. Demetriou}{equal, yyy}
\icmlauthor{Cynthia C. S. Liem}{yyy}
\end{icmlauthorlist}

@@

\icmlcorrespondingauthor{Firstname1 Lastname1}{first1.last1@xxx.edu}

@@

\subsection{Do neural networks get born with world models?}\label{example-deep-learning}

@@

%\subsection{Example: Large Language Model}\label{ex-llm}
\subsection{LLMs for economic sentiment prediction}\label{ex-llm}

@@

%\subsubsection{Stochastic Parrots AfterAll?}\label{stochastic-parrots-after-all}
\subsubsection{Sparks of economic understanding?}\label{stochastic-parrots-after-all}
%The results from the linear probes shown in Figure~\ref{fig-fomc} are certainly not unimpressive:
Even though \emph{FOMC-RoBERTa} (which is substantially smaller than the models tested in~\citet{gurnee2023languagev2} was not explicitly trained to uncover associations between central bank communications and the level of consumer prices, it appears that the model has distilled representations that can be used to predict inflation (although they certainly will not win any forecasting competitions). So, have we uncovered further evidence that LLMs ``aren't mere stochastic
parrots''? Has \emph{FOMC-RoBERTa} developed an intrinsic `understanding'
of the economy just by `reading' central bank communications? Should the FOMC use this LLM to direct their forward guidance?

% ICML sentence: Thus, can economists readily adopt FOMC-RoBERTa as a domain-relevant tool?

@@

\section{On human proneness to over-interpretation}
Linear probes and related tools from mechanistic interpretability were proposed in the context of monitoring models and diagnosing potential problems \citep{alain2018understanding}. Favorable outcomes from probes merely indicate that the model ``has learned information relevant for the property [of interest]" \citep{belinkov2021probing}. Our examples demonstrate, that this is achievable even for small models, while these models have certainly not developed an intrinsic ``understanding'' of the world. Thus, to draw conclusions about the emerging capabilities of AI models, we argue that more conservative and rigorous tests are needed.

>>>>

\begin{icmlauthorlist}
\icmlauthor{Patrick Altmeyer}{equal,yyy}
\icmlauthor{Andrew M. Demetriou}{equal,yyy}
\icmlauthor{Antony Bartlett}{yyy}
\icmlauthor{Cynthia C. S. Liem}{yyy}
\end{icmlauthorlist}

@@

\icmlcorrespondingauthor{Patrick Altmeyer}{patrick.altmeyer@tudelft.nl}

@@

\subsection{Are Neural Networks Born with World Models?}\label{example-deep-learning}

@@

%\subsection{Example: Large Language Model}\label{ex-llm}
\subsection{LLMs for Economic Sentiment Prediction}\label{ex-llm}

@@

%\subsubsection{Stochastic Parrots AfterAll?}\label{stochastic-parrots-after-all}
\subsubsection{Sparks of economic understanding?}\label{stochastic-parrots-after-all}
%The results from the linear probes shown in Figure~\ref{fig-fomc} are certainly not unimpressive:
Even though \emph{FOMC-RoBERTa} (which is substantially smaller than the models tested in~\citet{gurnee2023languagev2} was not explicitly trained to uncover associations between central bank communications and the level of consumer prices, it appears that the model has distilled representations that can be used to predict inflation (although they certainly will not win any forecasting competitions). So, have we uncovered further evidence that LLMs ``aren't mere stochastic
parrots''? Has \emph{FOMC-RoBERTa} developed an intrinsic `understanding'
of the economy just by `reading' central bank communications? Thus, can economists readily adopt FOMC-RoBERTa as a domain-relevant tool?

@@

\section{On Human Proneness to Over-Interpretation}
Linear probes and related tools from mechanistic interpretability were proposed in the context of monitoring models and diagnosing potential problems \citep{alain2018understanding}. Favorable outcomes from probes merely indicate that the model ``has learned information relevant for the property [of interest]" \citep{belinkov2021probing}. Our examples demonstrate, that this is achievable even for small models, while these models have certainly not developed an intrinsic ``understanding'' of the world. Thus, to draw conclusions about the emerging capabilities of AI models, we argue that more conservative and rigorous tests are needed.
